Evaluating a simplified forecast model in comparison to the European forecasting hub ensemble

### 2022-09-30 paper/paper.Rmd
add KS as a code reviewer
spell check

Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts over the last decade [@Reich2022-jo].
These approaches have been inspired by the experience of other domains, for example, climate modelling [@IPCCreport], where ensembles of both multiple models and from multiple teams have a long history of providng forecasts that are trusted by stakholders.

@@

The trend towards large-scale multi-team ensemble forecasting in infectious diseases has accelerated during the COVID-19 pandemic due to a pressing need for reliable forecasts and a perception that many publicly available forecasts were low quality. Over 2020 and 2021, teams established COVID-19 Forecasting Hubs covering the US [@Cramer2022-sp], Germany and Poland [@Bracher2021-wk], and Europe [@Sherratt2022-ue] (the latter two including authors of this study). All of these collaborations ensembled contributions from multiple independent teams using a similar approach and have shown that their ensemble forecasts outperform most individually contributed forecasts whilst remaining generally robust to outliers in reporting.
However, less progress has been made mitigating the downsides of these projects and improving access to the high quality and robust forecasts they seek to generate to geographies without coverage or for other infectious diseases.

In this work, we draw insights from ensemble forecasts produced and endorsed by these forecast hubs, as well as our own forecasting work, to propose and evaluate a "surrogate" forecast model.
This seeks to mitigate some of the downsides of recent COVID-19 ensembles forecasts by emulating their observed behaviour, using a minimal set of easily communicated and epidemiologically justifiable assumptions, and limited computational resources with an easily generalised implementation.
The primary aim of this approach is to help highlight the behaviour, and potential mechanisms behind this behaviour, of ensemble forecast widely considered the gold standard for COVID-19 forecasting.
In addition to this, we seek to provide a simple, and robust, forecasting system that can be easily reused by others, perhaps where a forecast hub ensemble is not available, and be used as a platform for future research both into improving understanding of ensemble behaviour and increased forecast performance.

In climate forecasting [@Castelletti2012-bs; @Edwards2021-lv; @Williamson2013-qx], as well as in other fields such as astrophysic [@Vernon2014-ov], emulation approaches have been used to circumvent resource requirement issues for complex models by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components.
However, within epidemiological modelling, we often want to use our models for applications other than forecasting (such as learning about underlying disease dynamics or developing scenarios) and, for ensemble models in particular, have fewer insights beyond forecast performance on which to base our mechanistic assumptions.
This means that, whilst non-parametric emulation has utility for allowing rapid exploration of the parameter space for epidemiological models [@Iskauskas2022-fz; @harles2022-ds], it may be less suitable for use in replicating epidemiological ensemble forecasts than the surrogate assumption driven framework we propose.

In order to understand the behaviour of the forecast hub ensembles we need to first explore the structure of the COVID-19 Forecast Hubs [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
These collaborations share a similar design with a central team running the hub, vetting forecasts, and producing the ensemble forecast as well as teams of independent forecast contributors who design their own forecast models and then use them to produce weekly forecast that they then submit to the central hub team. 
Each hub targets a range of metrics including test positive reported cases, repoorted deaths, and hospitalisations, has a specific geographic focus, and asks for weekly forecasts (using epidemiological weeks i.e. Sunday to Saturday) over a time horizon of a few weeks.
Typically, the ensemble forecast produced each week comprises at least 3 independent forecasts using a range of approaches and is produced using an unweighted median ensemble [@Sherratt2022-ue].
Observed data are available and updated daily, and most submitted forecasts use this dataset, along with potentially other sources of real-time information, to produce forecasts.
Here, we focus on reported cases and primarily on the European forecast hub but our observations hold, in our view, across COVID-19 forecast hubs and to a lesser degree targets.
In general, 4 main classes of forecast models are submitted, statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF; @Bosse2022-hm], and human insight based forecast models that may also include elements of other methods [@Bosse2022-hm].
Real-time evaluation has shown that each of these classes of models may perform well depending on the context and specific implementation of the forecast model [@Bosse2022-hm].

We have contributed a range of forecasts to these forecast collaborations, generally focussed on semi-mechanistic statistical methods, and human insight-based forecasts.
Our forecast submissions have not clearly over- or under- performed other forecasts submitted to the forecasting hubs (see `epiforecast` tagged models at [@EuroHub] and [@Bosse2022-hm; @Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue]).
The model-based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within the week, reporting periodicity [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
Based on our observations our forecasts have generally captured the current trend relatively well, but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns. 
Our previous methodology also requires significant computational resources, running for an hour on a Azure D v5-series 16-core machine, when producing forecasts for the European forecasting hub [@AzurePricing]. 
In our model-based forecasts, we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data whereas in our human insight models these were implicitly included.
We found that our human insight-based forecasts significantly outperformed our model-based forecasts on average and hypothesised that this may have been driven by including additional information not observed in the epidemiological data [@Bosse2022-hm].

Unlike our epidemiologically motivated forecast submissions, the hub ensemble forecasts were typically robust to daily reporting artefacts. They also demonstrated some ability to forecast future changes in trends that were not present in the observed data similarly to our human insight forecasts indicating the likely inclusion of either human insight, or assumptions about future interventions. 
In general, the ensemble forecasts were more auto-correlated and less reactive to changes from stable or reducing case incidence to increasing incidence. On the other hand, this also meant that the ensemble was less likely to adopt short-term changes in incidence and hence produced better long-term forecasts. Finally, the ensemble forecast tended to produce sharper forecasts and have a tendency towards under vs over predicting. There is some suggestion that this was partially driven by the ensemble forecast having non-exponential uncertainty, unlike many infectious disease forecasting models

Based on these insights, here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@Bosse2022-hm] with the aim of producing ensemble-like performance without sacrificing interpretability.
The first simplification we make is to model only weekly data, rather than using daily data and then aggregating. This mitigates the impact of daily reporting artefacts. It also serves to increase the auto-correlation of the forecasting model as there is an increased lag before changes in daily observations gain significant weight in the model.
This leads to the observed ensemble behaviour of being relatively auto-correlated and resistant to short-term changes in trend.

@@

We extracted forecasts and data on notified weekly COVID-19 cases from the European forecasting hub [@Sherratt2022-ue; @EuroHub] from the 15th of January 2022 to the 19th of July 2022 for the ensemble model (referred to as the `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth`).
We included all locations covered by the European forecasting hub which were 32 European countries, including all countries of the European Union and European Free Trade Area, and the United Kingdom.
Data on notified weekly cases was originally sourced from the Johns Hopkins University (JHU) curated data repository [@JHU].
We used the latest available observed data as of the 1st of September 2022 (commit `f6922c3e4bdcb055abcbba8e73472afacac4cf40` from [@EuroHubGitHub]).
Incidence was aggregated over the Morbidity and Mortality Weekly Report (MMWR) epidemiological week definition of Sunday through Saturday.
As observations are subject to revisions this means that the data used to produce forecasts for a given date may not reflect the data used for evaluation.
To account for this we followed the practice of the European forecasting hub project in excluding forecasts made using anomolous truth data in the week of the forecasts production and excluding forecasts for target weeks with anomolous data [@Sherratt2022-ue].
We defined anomolous data based on the implementation used by [@EuroHub] where a datapoint is considered anomolous if a future revision alters it by more than 5%.

@@

In addition to presenting the WIS for a subset of locations and the relative WIS for all locations, we also calculate and visualise the empirical coverage, which is the percentage of observed values within a given interval or below a given quantile, of both the surrogate and ensemble model for the 30%, 60%, and 90% prediction intervals and by quantile [@scoringutils]. 
We also calculate the bias [@Funk2019-vi] of both forecasting approaches, stratified by forecast horizon. This metric aims to capture the tendency for a forecast to under- or over-predict. It captures the average proportion of the mass of the forecast distribution that is above or below the true value (and so can range from -1 to 1) with an unbiased forecast having an average bias value of 0.
Lastly, we calculate and visualise the relative weighted interval score by quantile, stratified by forecast horizon, to assess the relative difference in performance across the predictive distribution.

@@

The model is implemented in `stan` [@stan] and `R` (`4.2.0`) [@R] as an extension of the baseline model from the `forecast.vocs` R package (`0.0.9.7000`)  [@forecast.vocs].
We note that our use of a Heaviside step function introduces a discontinuity to the posterior making it less suited for use with `stan`. Other model formulations without this feature would be more efficient and robust.
The `cmdstanr` R package (`0.5.2`) [@cmdstanr] is used for model fitting with 2 MCMC chains each having 1000 warm-up and 1000 sampling steps each [@cmdstanr].
`cmdstanr` surfaces several settings that trade off between sampling speed and the robustness of the approach.
Here we take a conservative approach, as the model fit is not manually inspected during real-time usage and due to the expected complexity of the posterior [@betancourt_2017], and set the adapt delta setting to 0.99, and the maximum tree depth setting to 15. 
For real-time usage, convergence was not assessed, but during model development, the Rhat diagnostic was used alongside feedback from `cmdstanr` about the number of divergent transitions and exceedance of the maximum tree depth [@cmdstanr]. During development, posterior predictions were also visually compared to observed data.

@@

To ensure the reproducibility of this analysis dependencies are managed using the `renv` R package (`0.14.0`) [@renv] and a Dockerfile file along with a built Docker image [@Boettiger:2015dw] (via GitHub Actions [@GitHubActionsAbout]) is provided in the code repository.
Weekly forecasts were made using `renv` and based GitHub Actions free tier as available in 2022 to ensure they require limited compute and that our implementation is independent of local resources facilitating democratised access.
The free GitHub Actions runner we used for all forecasts was Ubuntu `20.04` based with 2 cores (x86_64), 7 GB of RAM, and 14 GB of SSD space.
The code for this analysis can be found here: https://github.com/epiforecasts/simplified-forecaster-evaluation
The code for the forecasting model defined above along with the infrastructure required to forecast using GitHub Actions can be found here: https://github.com/seabbs/ecdc-weekly-growth-forecasts
Versions archived on Zenodo are available [@REF] and [@REF].

@@

We extracted forecasts starting from the 15th of January until the 19th of July 2022 for all countries covered by the European forecasting hub (nations of the European Union, European Free Trade Agreement, and the United Kingdom, making `r locations` unique locations).
In total `r nrow(metadata)` forecasts were made across all locations, with `r forecast_dates` unique forecast dates and `r models` independent forecast models (including the European hub baseline model).
Of these models, `r length(global_models)` forecasted in at least 30 locations including our original submission  (referred to as `epiforecasts-EpiNow2` by the hub), and our surrogate model.
Of the remaining models submitted `r length(local_models)` were submitted in only one location.
Single location models were clustered in a few locations, particularly Germany and Poland (likely due to the folding of the German/Poland forecasting hub into the European forecasting hub project [@Sherratt2022-ue]).
Italy was also an outlier with `r single_locs_models_per_loc[location == "IT", n]` models that submitted nowhere else.
`r nrow(partial_models)` models were submitted for between 3 and 30 locations and all these models varied the number of locations they submitted forecasts for over time, potentially indicating manual curation or models targeted at specific conditions.

@@

`r nrow(anomalies)` observations, stratified by week and location, were defined to be anomalous within the study period by the European forecast hub [@EuroHub]. 
Forecasts for these observations were excluded as were forecasts for forecast weeks where they were the latest available data.
Data anomalies were not randomly distributed with some locations being particlarly prone to data revisions including Lithuania (with `r anomalies[location == "LT", .N]` weeks with data anomalies), and Portugal (with `r anomalies[location == "PT", .N]` weeks with data anomalies).
Anomalies  were also not evenly distributed over time with a higher proportion occurring earlier in the study period (potentially due to our choice to extract data from the 1st of September which effectively truncated anomalies).
`r per_excluded`% of forecasts were excluded across all horizons due to anomalies in the observed data. Aggregated across horizons `r exclusions$per`% of forecasts included at least one week with anomalous data.

@@

In our example set of locations, the absolute performance of the ensemble and the surrogate model was visually similar on the log scale in all locations at short forecast horizons though this varied by location (Figure \@ref(fig:vis-forecasts) b).
On the natural scale the difference in performance was more marked, specially for periods of high incidence and at longer horizons (Figure \@ref(fig:vis-forecasts) a).
Performance was not homogeneous across our set of example locations with the surrogate model performing similarly to the ensemble in Slovakia whilst in the United Kingdom and Germany the surrogate model performed substantially worse for some forecast dates (Figure \@ref(fig:vis-forecasts)).
For both the ensemble and the surrogate, performance decreased as the forecast horizon increased with this being particularly noticable for the surrogate model during periods of high incidence.
In general, in the study period the ensemble appeared to be better able to forecast peak incidence. 
Both models forecast large reductions in incidence in Poland during May that did not occur whilst only the ensemble forecast spuriously forecast similar large reductions in Germany during June.
In comparison to the ensemble model the surrogate model appeared less likely to place weight on unfeasibly large reductions in incidence during periods of declining incidence but on other hand was more likely to forecast continuing increases in incidence (for example in February in Slovakia and Poland).

@@

Evaluating the ensemble and surrogate models using the WIS across all locations and forecast dates we found that the mean relative performance of the surrogate model was `r relative_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_wis_by_horizon[horizon == 4, ris]` at the four week horizon, indicating that the ensemble forecast outperformed the surrogate forecast for all horizons by at least 25% and that the relative performance of the surrogate model degraded as the forecast horizon increased  (Figure \@ref(fig:assemble-eval) c).
Much of this outperformance, especially at longer forecast horizons, was driven by a relatively small subset of forecasts with relative performance having a heavy tail (Figure \@ref(fig:assemble-eval) a). 
If we instead consider median relative performance (note this is not a proper scoring rule) we find that, relative to the ensemble, the surrogate scored `r relative_median_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_median_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_median_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_median_wis_by_horizon[horizon == 4, ris]` at the four week horizon.
This would suggest that an increasingly skewed score distribution as the forecast horizon increased is responsible for the increase in the mean relative score (Figure \@ref(fig:assemble-eval) a).
`r relative_wis_thresholds[, better_than]` of individual surrogate forecasts scored better than the comparable ensemble forecast, `r relative_wis_thresholds[, fifty_percent]` performed within 50% of the comparable ensemble forecast, and `r relative_wis_thresholds[, more_than_100_percent]` had a more than 100% worse WIS than the comparable ensemble forecast.

If we consider only the median point forecast, using the absolute error, we see that the ensemble forecast again outperformed the surrogate forecast (`r overall_ae_median[model == "Ensemble", round(Mean, digits = 1)]` vs. `r overall_ae_median[model == "Surrogate", round(Mean, digits = 1)]`). As when using the weighted interval score if we instead consider the median of the absolute error we see that the difference in performance has reduced indicating a similar skewed score distribution for point forecasts as for the whole predictive distribution (`r overall_ae_median[model == "Ensemble", round(Median, digits = 1)]` vs. `r overall_ae_median[model == "Surrogate", round(Median, digits = 1)]`). Across forecast horizons the same pattern of outperformance holds. However, the difference in relative performance was less than when the full probability distribution was accounted for with this becoming more marked as the forecast horizon increased (Figure \@ref(fig:assemble-eval) c).

@@

Overall the surrogate model was relatively well calibrated at the 30%, 60% and 90% prediction interval, though with a tendency to over cover, with empirical coverage of `r overall_coverage[range %in% "30% interval", Surrogate]`%, `r overall_coverage[range %in% "60% interval", Surrogate]`%, `r overall_coverage[range %in% "90% interval", Surrogate]`% respectively.
The ensemble model was less well calibrated, with a tendency to under cover with empirical coverage of `r overall_coverage[range %in% "30% interval", Ensemble]`%, `r overall_coverage[range %in% "60% interval", Ensemble]`%, `r overall_coverage[range %in% "90% interval", Ensemble]`% respectively (Figure \@ref(fig:plot-coverage) a).
When stratified by forecast horizon the ensemble forecast was best calibrated at the one-week forecast horizon, and then became progressively less well calibrated as the forecast horizon increased (Figure \@ref(fig:plot-coverage) a). 
In comparison, the surrogate forecast was less well calibrated than the ensemble forecast at the one-week forecast horizon with a tendency to have a larger empirical coverage than required  (Figure \@ref(fig:plot-coverage) a).
At longer horizons and narrower prediction intervals the surrogate forecast became better calibrated though with a tendency to under cover.
This was not the case for the 90% prediction interval where the surrogate model covered more than the expected interval, for all horizons, indicating forecasts were overly uncertain for this interval regardless of horizon.

Stratifying calibration by quantile and forecast horizon the ensemble forecast was overly certain at all horizons for quantiles larger than the median whilst being comparably well calibrated for intervals below the median (Figure \@ref(fig:plot-coverage) b).
This behaviour became more prominant as forecast horizon increased.
In contrast, the surrogate forecast was generally equally well calibrated across horizons with a tendency to over cover for intervals above the median.
At longer horizons, however, quantiles below the median became liable to under cover.

>>>>

Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts over the last decade [@Reich2022-jo].
These approaches have been inspired by the experience of other domains, for example, climate modelling [@IPCCreport], where ensembles of both multiple models and from multiple teams have a long history of providng forecasts that are trusted by stakeholders.

@@

The trend towards large-scale multi-team ensemble forecasting in infectious diseases has accelerated during the COVID-19 pandemic due to a pressing need for reliable forecasts and a perception that many publicly available forecasts were low quality. Over 2020 and 2021, teams established COVID-19 Forecasting Hubs covering the US [@Cramer2022-sp], Germany and Poland [@Bracher2021-wk], and Europe [@Sherratt2022-ue] (the latter two including authors of this study). All of these collaborations ensembled contributions from multiple independent teams using a similar approach and have shown that their ensemble forecasts outperform most individually contributed forecasts whilst remaining generally robust to outliers in reporting.
However, less progress has been made in mitigating the downsides of these projects and improving access to the high-quality and robust forecasts they seek to generate to geographies without coverage or for other infectious diseases.

In this work, we draw insights from ensemble forecasts produced and endorsed by these forecast hubs, as well as our forecasting work, to propose and evaluate a "surrogate" forecast model.
This seeks to mitigate some of the downsides of recent COVID-19 ensembles forecasts by emulating their observed behaviour, using a minimal set of easily communicated and epidemiologically justifiable assumptions, and limited computational resources with an easily generalised implementation.
The primary aim of this approach is to help highlight the behaviour, and potential mechanisms behind this behaviour, of ensemble forecast widely considered the gold standard for COVID-19 forecasting.
In addition to this, we seek to provide a simple, and robust, forecasting system that can be easily reused by others, perhaps where a forecast hub ensemble is not available, and be used as a platform for future research both into improving understanding of ensemble behaviour and increased forecast performance.

In climate forecasting [@Castelletti2012-bs; @Edwards2021-lv; @Williamson2013-qx], as well as in other fields such as astrophysics [@Vernon2014-ov], emulation approaches have been used to circumvent resource requirement issues for complex models by training a simplified model, usually, a non-parametric statistical model, to replicate the behaviour of either the entire model or sub-components.
However, within epidemiological modelling, we often want to use our models for applications other than forecasting (such as learning about underlying disease dynamics or developing scenarios) and, for ensemble models, in particular, have fewer insights beyond forecast performance on which to base our mechanistic assumptions.
This means that, whilst non-parametric emulation has utility for allowing rapid exploration of the parameter space for epidemiological models [@Iskauskas2022-fz; @harles2022-ds], it may be less suitable for use in replicating epidemiological ensemble forecasts than the surrogate assumption driven framework we propose.

To understand the behaviour of the forecast hub ensembles we need to first explore the structure of the COVID-19 Forecast Hubs [@Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue].
These collaborations share a similar design with a central team running the hub, vetting forecasts, and producing the ensemble forecast as well as teams of independent forecast contributors who design their forecast models and then use them to produce a weekly forecast that they then submit to the central hub team. 
Each hub targets a range of metrics including test-positive reported cases, reported deaths, and hospitalisations has a specific geographic focus and asks for weekly forecasts (using epidemiological weeks i.e. Sunday to Saturday) over a time horizon of a few weeks.
Typically, the ensemble forecast produced each week comprises at least 3 independent forecasts using a range of approaches and is produced using an unweighted median ensemble [@Sherratt2022-ue].
Observed data are available and updated daily, and most submitted forecasts use this dataset, along with potentially other sources of real-time information, to produce forecasts.
Here, we focus on reported cases and primarily on the European forecast hub but our observations hold, in our view, across COVID-19 forecast hubs and to a lesser degree targets.
In general, 4 main classes of forecast models are submitted, statistical forecasting models such as ARIMA models [@REF], mechanistic forecasting models based on the compartmental modelling framework and its generalisations [@REF], semi-mechanistic approaches that blend both of these approaches [@REF; @Bosse2022-hm], and human insight based forecast models that may also include elements of other methods [@Bosse2022-hm].
Real-time evaluation has shown that each of these classes of models may perform well depending on the context and specific implementation of the forecast model [@Bosse2022-hm].

We have contributed a range of forecasts to these forecast collaborations, generally focused on semi-mechanistic statistical methods and human insight-based forecasts.
Our forecast submissions have not over- or under-performed other forecasts submitted to the forecasting hubs (see `epiforecast` tagged models at [@EuroHub] and [@Bosse2022-hm; @Cramer2022-sp; @Bracher2021-wk; @Sherratt2022-ue]).
The model-based forecasts we have contributed have focussed on trying to carefully model the underlying infectious disease dynamics from infection through to symptom onset, and test positivity using non-exponential delay distributions whilst also attempting to model the complexity of daily, within the week, reporting periodicity [@Bosse2022-hm; @Abbott2020-iy; @EpiNow2].
Based on our observations our forecasts have generally captured the current trend relatively well but have not been robust to reporting issues such as large outliers in reporting and changes to reporting patterns. 
Our previous methodology also requires significant computational resources, running for an hour on a Azure D v5-series 16-core machine, when producing forecasts for the European forecasting hub [@AzurePricing]. 
In our model-based forecasts, we did not attempt to capture potential future interventions or known interventions not currently observed in the epidemiological data whereas in our human insight models these were implicitly included.
We found that our human insight-based forecasts significantly outperformed our model-based forecasts on average and hypothesised that this may have been driven by including additional information not observed in the epidemiological data [@Bosse2022-hm].

Unlike our epidemiologically motivated forecast submissions, the hub ensemble forecasts were typically robust to daily reporting artefacts. They also demonstrated some ability to forecast future changes in trends that were not present in the observed data similarly to our human insight forecasts indicating the likely inclusion of either human insight, or assumptions about future interventions. 
In general, the ensemble forecasts were more auto-correlated and less reactive to changes from stable or reducing case incidence to increasing incidence. On the other hand, this also meant that the ensemble was less likely to adopt short-term changes in incidence and hence produced better long-term forecasts. Finally, the ensemble forecast tended to produce sharper forecasts and have a tendency toward under vs over predicting. There is some suggestion that this was partially driven by the ensemble forecast having non-exponential uncertainty, unlike many infectious disease forecasting models

Based on these insights, here we define a model with similar, but simplified, epidemiological characteristics to our previous approaches to model-based forecasting [@Bosse2022-hm] to produce an ensemble-like performance without sacrificing interpretability.
The first simplification we make is to model only weekly data, rather than using daily data and then aggregating. This mitigates the impact of daily reporting artefacts. It also serves to increase the auto-correlation of the forecasting model as there is an increased lag before changes in daily observations gain significant weight in the model.
This leads to the observed ensemble behaviour of being relatively auto-correlated and resistant to short-term changes in trend.

@@

We extracted forecasts and data on notified weekly COVID-19 cases from the European forecasting hub [@Sherratt2022-ue; @EuroHub] from the 15th of January 2022 to the 19th of July 2022 for the ensemble model (referred to as the `EuroCOVIDhub-ensemble` by the hub team) and the surrogate model (submitted as `epiforecasts-weeklygrowth`).
We included all locations covered by the European forecasting hub which were 32 European countries, including all countries of the European Union and European Free Trade Area, and the United Kingdom.
Data on notified weekly cases was originally sourced from the Johns Hopkins University (JHU) curated data repository [@JHU].
We used the latest available observed data as of the 1st of September 2022 (commit `f6922c3e4bdcb055abcbba8e73472afacac4cf40` from [@EuroHubGitHub]).
Incidence was aggregated over the Morbidity and Mortality Weekly Report (MMWR) epidemiological week definition of Sunday through Saturday.
As observations are subject to revisions this means that the data used to produce forecasts for a given date may not reflect the data used for evaluation.
To account for this we followed the practice of the European forecasting hub project in excluding forecasts made using anomalous truth data in the week of the forecasts production and excluding forecasts for target weeks with anomalous data [@Sherratt2022-ue].
We defined anomalous data based on the implementation used by [@EuroHub] where a datapoint is considered anomalous if a future revision alters it by more than 5%.

@@

In addition to presenting the WIS for a subset of locations and the relative WIS for all locations, we also calculate and visualise the empirical coverage, which is the percentage of observed values within a given interval or below a given quantile, of both the surrogate and ensemble model for the 30%, 60%, and 90% prediction intervals and by quantile [@scoringutils]. 
We also calculate the bias [@Funk2019-vi] of both forecasting approaches, stratified by forecast horizon. This metric aims to capture the tendency for a forecast to under or over-predict. It captures the average proportion of the mass of the forecast distribution that is above or below the true value (and so can range from -1 to 1) with an unbiased forecast having an average bias value of 0.
Lastly, we calculate and visualise the relative weighted interval score by quantile, stratified by forecast horizon, to assess the relative difference in performance across the predictive distribution.

@@

The model is implemented in `stan` [@stan] and `R` (`4.2.0`) [@R] as an extension of the baseline model from the `forecast.vocs` R package (`0.0.9.7000`)  [@forecast.vocs].
We note that our use of a Heaviside step function introduces a discontinuity to the posterior making it less suited for use with `stan`. Other model formulations without this feature would be more efficient and robust.
The `cmdstanr` R package (`0.5.2`) [@cmdstanr] is used for model fitting with 2 MCMC chains each having 1000 warm-up and 1000 sampling steps each [@cmdstanr].
`cmdstanr` surfaces several settings that trade-off between sampling speed and the robustness of the approach.
Here we take a conservative approach, as the model fit is not manually inspected during real-time usage and due to the expected complexity of the posterior [@betancourt_2017], and set the adapt delta setting to 0.99, and the maximum tree depth setting to 15. 
For real-time usage, convergence was not assessed, but during model development, the Rhat diagnostic was used alongside feedback from `cmdstanr` about the number of divergent transitions and exceedance of the maximum tree depth [@cmdstanr]. During development, posterior predictions were also visually compared to observed data.

@@

To ensure the reproducibility of this analysis dependencies are managed using the `renv` R package (`0.14.0`) [@renv] and a Dockerfile file along with a built Docker image [@Boettiger:2015dw] (via GitHub Actions [@GitHubActionsAbout]) is provided in the code repository.
Weekly forecasts were made using `renv` and based on GitHub Actions free tier as available in 2022 to ensure they require limited compute and that our implementation is independent of local resources facilitating democratised access.
The free GitHub Actions runner we used for all forecasts was Ubuntu `20.04` based with 2 cores (x86_64), 7 GB of RAM, and 14 GB of SSD space.
The code for this analysis can be found here: https://github.com/epiforecasts/simplified-forecaster-evaluation
The code for the forecasting model defined above along with the infrastructure required to forecast using GitHub Actions can be found here: https://github.com/seabbs/ecdc-weekly-growth-forecasts
Versions archived on Zenodo are available [@REF] and [@REF].

@@

We extracted forecasts starting from the 15th of January until the 19th of July 2022 for all countries covered by the European forecasting hub (nations of the European Union, the European Free Trade Agreement, and the United Kingdom, making `r locations` unique locations).
In total `r nrow(metadata)` forecasts were made across all locations, with `r forecast_dates` unique forecast dates and `r models` independent forecast models (including the European hub baseline model).
Of these models, `r length(global_models)` forecasted in at least 30 locations including our original submission  (referred to as `epiforecasts-EpiNow2` by the hub), and our surrogate model.
Of the remaining models submitted `r length(local_models)` were submitted in only one location.
Single location models were clustered in a few locations, particularly in Germany and Poland (likely due to the folding of the German/Poland forecasting hub into the European forecasting hub project [@Sherratt2022-ue]).
Italy was also an outlier with `r single_locs_models_per_loc[location == "IT", n]` models that submitted nowhere else.
`r nrow(partial_models)` models were submitted for between 3 and 30 locations and all these models varied the number of locations they submitted forecasts for over time, potentially indicating manual curation or models targeted at specific conditions.

@@

`r nrow(anomalies)` observations, stratified by week and location, were defined to be anomalous within the study period by the European forecast hub [@EuroHub]. 
Forecasts for these observations were excluded as were forecasts for forecast weeks where they were the latest available data.
Data anomalies were not randomly distributed with some locations being particularly prone to data revisions including Lithuania (with `r anomalies[location == "LT", .N]` weeks with data anomalies), and Portugal (with `r anomalies[location == "PT", .N]` weeks with data anomalies).
Anomalies were also not evenly distributed over time with a higher proportion occurring earlier in the study period (potentially due to our choice to extract data from the 1st of September which effectively truncated anomalies).
`r per_excluded`% of forecasts were excluded across all horizons due to anomalies in the observed data. Aggregated across horizons `r exclusions$per`% of forecasts included at least one week with anomalous data.

@@

In our example set of locations, the absolute performance of the ensemble and the surrogate model was visually similar on the log scale in all locations at short forecast horizons though this varied by location (Figure \@ref(fig:vis-forecasts) b).
On the natural scale the difference in performance was more marked, especially for periods of high incidence and at longer horizons (Figure \@ref(fig:vis-forecasts) a).
Performance was not homogeneous across our set of example locations with the surrogate model performing similarly to the ensemble in Slovakia whilst in the United Kingdom and Germany the surrogate model performed substantially worse for some forecast dates (Figure \@ref(fig:vis-forecasts)).
For both the ensemble and the surrogate, performance decreased as the forecast horizon increased with this being particularly noticeable for the surrogate model during periods of high incidence.
In general, in the study period, the ensemble appeared to be better able to forecast peak incidence. 
Both models forecast large reductions in incidence in Poland during May that did not occur whilst only the ensemble forecast spuriously forecast similar large reductions in Germany during June.
In comparison to the ensemble model the surrogate model appeared less likely to place weight on unfeasibly large reductions in incidence during periods of declining incidence but on other hand was more likely to forecast continuing increases in incidence (for example in February in Slovakia and Poland).

@@

Evaluating the ensemble and surrogate models using the WIS across all locations and forecast dates we found that the mean relative performance of the surrogate model was `r relative_wis_by_horizon[horizon == 1, ris]` at the one-week horizon, `r relative_wis_by_horizon[horizon == 2, ris]` at the two-week horizon, `r relative_wis_by_horizon[horizon == 3, ris]` at the three-week horizon, and `r relative_wis_by_horizon[horizon == 4, ris]` at the four-week horizon, indicating that the ensemble forecast outperformed the surrogate forecast for all horizons by at least 25% and that the relative performance of the surrogate model degraded as the forecast horizon increased  (Figure \@ref(fig:assemble-eval) c).
Much of this outperformance, especially at longer forecast horizons, was driven by a relatively small subset of forecasts with relative performance having a heavy tail (Figure \@ref(fig:assemble-eval) a). 
If we instead consider median relative performance (note this is not a proper scoring rule) we find that, relative to the ensemble, the surrogate scored `r relative_median_wis_by_horizon[horizon == 1, ris]` at the one week horizon, `r relative_median_wis_by_horizon[horizon == 2, ris]` at the two week horizon, `r relative_median_wis_by_horizon[horizon == 3, ris]` at the three week horizon, and `r relative_median_wis_by_horizon[horizon == 4, ris]` at the four week horizon.
This would suggest that an increasingly skewed score distribution as the forecast horizon increased is responsible for the increase in the mean relative score (Figure \@ref(fig:assemble-eval) a).
`r relative_wis_thresholds[, better_than]` of individual surrogate forecasts scored better than the comparable ensemble forecast, `r relative_wis_thresholds[, fifty_percent]` performed within 50% of the comparable ensemble forecast, and `r relative_wis_thresholds[, more_than_100_percent]` had a more than 100% worse WIS than the comparable ensemble forecast.

If we consider only the median point forecast, using the absolute error, we see that the ensemble forecast again outperformed the surrogate forecast (`r overall_ae_median[model == "Ensemble", round(Mean, digits = 1)]` vs. `r overall_ae_median[model == "Surrogate", round(Mean, digits = 1)]`). When using the weighted interval score if we instead consider the median of the absolute error we see that the difference in performance has reduced indicating a similar skewed score distribution for point forecasts as for the whole predictive distribution (`r overall_ae_median[model == "Ensemble", round(Median, digits = 1)]` vs. `r overall_ae_median[model == "Surrogate", round(Median, digits = 1)]`). Across forecast horizons the same pattern of outperformance holds. However, the difference in relative performance was less than when the full probability distribution was accounted for with this becoming more marked as the forecast horizon increased (Figure \@ref(fig:assemble-eval) c).

@@

Overall the surrogate model was relatively well calibrated at the 30%, 60% and 90% prediction interval, though with a tendency to over cover, with empirical coverage of `r overall_coverage[range %in% "30% interval", Surrogate]`%, `r overall_coverage[range %in% "60% interval", Surrogate]`%, `r overall_coverage[range %in% "90% interval", Surrogate]`% respectively.
The ensemble model was less well calibrated, with a tendency to under cover with empirical coverage of `r overall_coverage[range %in% "30% interval", Ensemble]`%, `r overall_coverage[range %in% "60% interval", Ensemble]`%, `r overall_coverage[range %in% "90% interval", Ensemble]`% respectively (Figure \@ref(fig:plot-coverage) a).
When stratified by forecast horizon the ensemble forecast was best calibrated at the one-week forecast horizon, and then became progressively less well calibrated as the forecast horizon increased (Figure \@ref(fig:plot-coverage) a). 
In comparison, the surrogate forecast was less well calibrated than the ensemble forecast at the one-week forecast horizon with a tendency to have a larger empirical coverage than required  (Figure \@ref(fig:plot-coverage) a).
At longer horizons and narrower prediction intervals, the surrogate forecast became better calibrated though with a tendency to under cover.
This was not the case for the 90% prediction interval where the surrogate model covered more than the expected interval, for all horizons, indicating forecasts were overly uncertain for this interval regardless of the horizon.

Stratifying calibration by quantile and forecast horizon the ensemble forecast was overly certain at all horizons for quantiles larger than the median whilst being comparably well calibrated for intervals below the median (Figure \@ref(fig:plot-coverage) b).
This behaviour became more prominent as the forecast horizon increased.
In contrast, the surrogate forecast was generally equally well calibrated across horizons with a tendency to over cover for intervals above the median.
At longer horizons, however, quantiles below the median became liable to under cover.
